\chapter{Introduction}

\vspace{0.5cm}

As seen in \ref{section:lojban-grammars}, the existing formal grammars for Lojban are extremely complete.
They cover the entirety of the language, both in its morphology and syntax/semantics, which means they can even validate words that don't exist yet
as long as they fit the language's rules (as mentioned in \ref{subsec:brief-intro-to-lojban}, an algorithm for "word creation" exists).
However, this completion makes these formal grammars extremely complex to read for a beginner, and are simply tools to be used by machines. \newline

As an example, here is an extract of the first few "root nodes" of GRAMMAR NAME: \newline

ADD EXTRACT OF GRAMMAR HERE \newline

What if there existed a formal grammar which people could read along, while learning Lojban, in order to gather a deeper understanding of the language?\newline

For any other language, using a formal grammar file as a learning companion would be a very obscure method.
However, Lojban being a language that can be parsed just as a programming language would, it would be very interesting to build a formal grammar
which would help a prospective student understand the intrinsic rules behind the language's grammar. \newline

Of course, improving the reading comprehension of a formal grammar reduces the complete aspect of it:

\begin{itemize}
  \item It is almost impossible to address all rules of the grammar, as some of them are extremely complex;
  \item Ignoring the morphology of words means that unknown but valid words won't be accepted;
\end{itemize}

However, as a beginner's tool, such a grammar could really improve the steep learning curve of Lojban, in my opinion.\newline

This chapter will explore the main reference for Lojban (\cite{cowan1997complete}), in its 2016 revised version, and write a simplified parser for a subset of the language.

% https://en.wikipedia.org/wiki/Extended_Backus%E2%80%93Naur_form

% https://en.wikipedia.org/wiki/Parsing_expression_grammar