\chapter{Thesis Work Plan}

\section{Deadlines}
\label{sec:deadlines}

The following broad outline is what I imagine, if possible, to be the list of big
deadlines for the academic year in which the thesis will be written:

\begin{itemize}
    \vspace{-0.4cm}
    \item \textbf{Status meetings:} Depending on the availability of the thesis director,
    I'd suggest holding these meetings (or emails) every month and a half to discuss advances
    on the thesis. For example, meetings in mid-October 2022, end of November 2022, mid-January 2023,
    end of February 2023 and mid-April 2023.
    \item \textbf{Final delivery:} Around May 2023.
    \item \textbf{Final presentation:} Around June 2023.
\end{itemize}

\vspace{-0.8cm}

\section{Organizing work}
\label{sec:organizing_work}

Given my background in IT, and as a software developer by profession, I would like
to organise my work for this Master Thesis as I'm used to: like a software project.
I'm planning to version the thesis \LaTeX \ document in a version control system
(most probably Git using Github). This allows to backup the document gradually,
as well as providing a better method to track the evolution of the document.\newline

I am also envisioning to write a "Changelog" file during the elaboration of the
thesis, and presenting it during the "status meetings" with the thesis director, as described above.
This allows for a better understanding of what changed between two versions of the thesis,
instead of having to read it in its entirety each time.
